---
title: "Metrics"
author: 'Author: L Pollock'
date: 'Last update: "`r Sys.Date()`"'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Metrics - Biotic Interactions

Biotic Interactions calculated from meta-web- who eats whom, binary interactions

Definitions of metawebs: (all are Adult and juvenille interactions, no eggs)
Metaweb1 - Potential (no co-occurrence/habitat requirement)
Metaweb2 - P-P must occur in at least 1 pixel together
Metaweb4 - GloBI interactions

For each species in each site, the following metrics are calculated:
-the number of prey species
-competition (see metrics)

Competition Metrics:
RA - richness of competitors (competitors=a species that shares at least one prey species)
R25 - richness of competitors (competitors=a species that shares at least 25% of its total prey species)
R90 - richness of strong competitors (competitors=a species that shares at least 90% of its total prey species)



Scenarios:
Run 1 - Metaweb1 + R25
Run 2 - Metaweb2 + R25
Run 3 - Metaweb2 + R90
Rub 4 - Metaweb2 + RA
Run 5 - Metaweb4 + RA

Notation is row.by.column
  

Set-up
```{r, eval=FALSE}
library(Matrix)

#site by species matrix
load("InputData/SiteBySpecies21Feb.Rdata")
#metaweb - adult-only interactions - amphibians have ZERO prey
#load("~/Documents/Manuscripts/EuropeanFoodWebCons/manuscript/Manuscript/Analyses3October18/MetaWebFromJ/BARMbinADULT.RData")

#metaweb - adult plus juvenille, no eggs
load("InputData/BARMbinWTeggs.RData")

#web from GloBI  ####  note:Predators are in columns
globi <- read.csv('InputData/globi_species_Dom.csv')
rownames(globi) <- globi[,1]
globi <- globi[-1]
globi2 <- t(globi)
rm(globi)

```

Metaweb 1 - Potential (no filters, only who 'could' eat whom)
```{r, eval=FALSE}

pred.by.prey <- BARM.binary
rm(BARM.binary)

PdPy <- pred.by.prey[intersect(colnames(site.by.species),rownames(pred.by.prey)),
                             intersect(colnames(site.by.species),colnames(pred.by.prey))]
PdPy[is.na(PdPy)]<-0
rm(pred.by.prey)

pred.by.prey <- Matrix(PdPy)
rm(PdPy)
if (sum(colSums(pred.by.prey))!=68390) {print("warning:check inputs")}

setwd("~/Documents/Manuscripts/EuropeanFoodWebCons/manuscript/Manuscript/Analyses21Feb19")
save(pred.by.prey,file="pred.by.prey.Metaweb1.Rdata")

rm(pred.by.prey)
```


Metaweb 2 - "Realized" - Potential interactions from MW 1, but also must co-occur in one or more pixels
```{r, eval=FALSE}
load("pred.by.prey.Metaweb1.Rdata")

if (any (!colnames(pred.by.prey) %in% colnames(site.by.species))) {print("warning:name mismatch")}

species.by.site <- t(site.by.species)

#diagonal=total pixel occurrences for each species, 
#off diagonal=joint probabilities (the intersection or the cells both occur in)
shared.sites.per.species <- species.by.site %*% site.by.species
#do sp co-occur or not
shared.sites.per.species[shared.sites.per.species > 1] <- 1

#scalar multiply for only pred-prey relationships that co-occur at least once
### 68390 pairs of pred/prey of potential, and 47061 of 'realized'
pred.by.prey <- pred.by.prey * shared.sites.per.species

if (sum(pred.by.prey)!=45809) {print("warning:wrong n. inters")}
save(pred.by.prey, file="pred.by.prey.Metaweb2Realized.Rdata")

```



Metaweb 4 - GloBI interactions
```{r, eval=FALSE}
load("pred.by.prey.Metaweb4Globi.Rdata")
pred.by.prey <- globi2
rm(globi2)

PdPy <- pred.by.prey[intersect(colnames(site.by.species),rownames(pred.by.prey)),
                             intersect(colnames(site.by.species),colnames(pred.by.prey))]
PdPy[is.na(PdPy)]<-0
rm(pred.by.prey)

pred.by.prey <- as.matrix(PdPy)
rm(PdPy)
if (sum(colSums(pred.by.prey))!=301) {print("warning:check inputs")}

save(pred.by.prey,file="pred.by.prey.Metaweb4Globi.Rdata")


#check which species have preds and prey
sp <- cbind(colSums(pred.by.prey),rowSums(pred.by.prey))
sp[sp>0] <- 1
spsum <- sp[,1] + sp[,2]
namz <- names(spsum[which(spsum %in% 2)]) #only those with pred and prey 

save(namz,file="MW4Globispecieslist.Rdata")

rm(pred.by.prey)
```



Calculate species-level metrics for main scenario (Metaweb 2: R25)

Set-up
```{r, eval=FALSE}

library(Matrix)

#site by species matrix
load("SiteBySpecies21Feb.Rdata")
#metaweb
load("pred.by.prey.Metaweb2Realized.Rdata")

if (any( colnames(site.by.species) %in% colnames(pred.by.prey)==FALSE)) {print("mismatched names")}


#set up list of stats to report
s <- c("RangeSize","meanSRsp","Nprey","Ncomp","RSprey","RScomp")
SpWebStats <- vector("list", length(s)) 
names(SpWebStats) <- s

#prep
prey.by.pred <- t(pred.by.prey)
species.by.site <- t(site.by.species)

#range size of each species
site.no.occur <- colSums(site.by.species)
SpWebStats[["RangeSize"]] <- site.no.occur

#ave species richness across cells occupied by species i
SR <- colSums(species.by.site)
spSR <- species.by.site*SR
totspSR <- rowSums(spSR)
meanspSR <- totspSR/site.no.occur
SpWebStats[["meanSRsp"]] <- meanspSR


#number of prey per predator (potential prey)
TotalNuPreySpeciesPerPred <- colSums(prey.by.pred)
SpWebStats[["Nprey"]] <- TotalNuPreySpeciesPerPred

#ave prey richness across cells occupied by predator i (realized prey across the landscape)
site.by.prey <- site.by.species %*% prey.by.pred
SpWebStats[["RZprey"]] <- colMeans(site.by.prey)

# mean number of occurrences across prey for each predator (average range size of prey for each pred)
prey.preyrichifpred <- prey.by.pred*site.no.occur
totalrichPreyPerPredSp <- colSums(prey.preyrichifpred)
meanRangSizPreyPerPredSp <- totalrichPreyPerPredSp / TotalNuPreySpeciesPerPred
SpWebStats[["RSprey"]] <- meanRangSizPreyPerPredSp

## competitors---------------
ShPrey <- pred.by.prey %*% prey.by.pred

#share 25% of prey
ShPrey <- pred.by.prey %*% prey.by.pred
cs <- ShPrey/colSums(prey.by.pred)
cs[cs<0.25] <- 0
cs[cs>=0.25] <- 1
cs[is.na(cs)] <- 0 
ShPrey <- t(cs)
diag(ShPrey) <- 0

#ave competitor richness across cells occupied by predator i (realized competitors across the landscape)
site.by.comp <- site.by.species %*% ShPrey
SpWebStats[["RZcomp"]] <- colMeans(site.by.comp)

#number of competitors
TotalNuCompetitorsPertarget <- colSums(ShPrey)
SpWebStats[["Ncomp"]] <- TotalNuCompetitorsPertarget

# mean number of occurrences across competitor for each target
sh <- ShPrey*site.no.occur
richsh <- colSums(sh)
meanRangSizCompetitorPerTargetSp <- richsh / TotalNuCompetitorsPertarget
SpWebStats[["RScomp"]] <- meanRangSizCompetitorPerTargetSp


save(SpWebStats,file = "SpeciesLevelWebStatisticsMW2R25.Rdata")

```


Set-up
```{r, eval=FALSE}

load("SpeciesLevelWebStatistics.Rdata")
load("MW2specieslist.Rdata")


t <- do.call("cbind",SpWebStats)

t.sp <- data.frame(t[which( rownames(t) %in% new.species),])
t.sp$group <- as.factor(gsub("[0-9]*","",rownames(t.sp)))

lst <- list("RangeSize","RSprey","RScomp")
tsp2 <- sapply(lst, function(x) round((t.sp[,x] / 1000),0) )
colnames(tsp2) <- unlist(lst)

t.sp <- cbind(t.sp[,c(9,2:4,7:8)],tsp2)


BirdCol <- "#7570b3"
MamCol <- "#d95f02"
AmphCol <- "#e7298a"
ReptCol <- "#1b9e77"


#function to make boxplots
boxplotz <- function(dat,col,labels,ylim) {
    for (x in col) {
    boxplot(dat[,x] ~ dat$group,ylab="",xlab="",col=c(AmphCol,BirdCol,MamCol,ReptCol),las=1,cex.axis=1.5,
            cex.lab=1.5,ylim=ylim)
    mtext(labels[x-1],side=2,padj=-4)
    }
}


labz <- c("MeanRichness","#Prey","#Competitors","Mean Prey Richness","Mean Comp. Richness","Range Size (1k pixels)","Range Size - Prey","Range Size - Comp")

pdf(file="Figure1stats.pdf",width=7,height=10)
par(mfrow=c(3,2))

boxplotz(t.sp,col=3:4,labz,ylim=c(0,300))
boxplotz(t.sp,col=5:6,labz,ylim=c(0,60))
boxplotz(t.sp,col=8:9,labz,ylim=c(0,320))

dev.off()

```



Set-up
```{r, eval=FALSE}

## combine both sets of species for appendix..


b <- merge(a,t.sp,by.x='Row.names',by.y=0,all=T)

#save(b,file="BspeciesCompare.Rdata")

load("BspeciesCompare.Rdata")


boxplotz2 <- function(dat,columns) {
    for (x in 1:length(columns)) {
    boxplot(dat[,columns[x]] ~ dat$group,ylab="",xlab="",col=c(AmphCol,BirdCol,MamCol,ReptCol),las=1,cex.axis=1.5,cex.lab=1.5)
    mtext(columns[x],side=2,padj=-4)
    }
}

# boxplots
pdf(file="Figure1statsContinuedOtherMetrics.pdf",width=12,height=8)

par(mfrow=c(2,5))

l <- c('prey','compRA','compACra','TL','Eig.cent','RangeSize','meanSRsp')

boxplotz2(b,l)

dev.off()

rm(t.sp,pred.by.prey)

```





Calculate species level metrics for GloBI interactions

Set-up
```{r, eval=FALSE}

library(Matrix)

#site by species matrix
load("~/Dropbox/Manuscripts/EuropeanFoodWebCons/manuscript/Manuscript/Analyses21Feb19/SiteBySpecies21Feb.Rdata")



#metaweb
#web from GloBI
load("pred.by.prey.Metaweb4Globi.Rdata")

if (any( colnames(site.by.species) %in% colnames(pred.by.prey)==FALSE)) {print("mismatched names")}


 
#set up list
s <- c("RangeSize","meanSRsp","Nprey","Npred","Ncomp","Napcomp","RSprey","RSpred","RScomp","RSapcomp")
SpWebStats <- vector("list", length(s)) 
names(SpWebStats) <- s

#prep
prey.by.pred <- t(pred.by.prey)
species.by.site <- t(site.by.species)

#range size of each species
site.no.occur <- colSums(site.by.species)
SpWebStats[["RangeSize"]] <- site.no.occur

#ave species richness across cells occupied by species i
SR <- colSums(species.by.site)
spSR <- species.by.site*SR
totspSR <- rowSums(spSR)
meanspSR <- totspSR/site.no.occur
SpWebStats[["meanSRsp"]] <- meanspSR

#number of prey per predator
TotalNuPreySpeciesPerPred <- colSums(prey.by.pred)
SpWebStats[["Nprey"]] <- TotalNuPreySpeciesPerPred

#number of predator per prey species
TotalNuPredSpeciesPerPrey <- colSums(pred.by.prey)
SpWebStats[["Npred"]] <- TotalNuPredSpeciesPerPrey
  
# mean number of occurrences across prey for each predator
prey.preyrichifpred <- prey.by.pred*site.no.occur
totalrichPreyPerPredSp <- colSums(prey.preyrichifpred)
meanRangSizPreyPerPredSp <- totalrichPreyPerPredSp / TotalNuPreySpeciesPerPred
SpWebStats[["RSprey"]] <- meanRangSizPreyPerPredSp

# mean number of occurrences across predator for each prey
pred.predrichifprey <- pred.by.prey*site.no.occur
totalrichPredPerPreySp <- colSums(pred.predrichifprey)
meanRangSizPredPerPreySp <- totalrichPredPerPreySp / TotalNuPredSpeciesPerPrey
SpWebStats[["RSpred"]] <- meanRangSizPredPerPreySp


## competitors---------------
ShPrey <- pred.by.prey %*% prey.by.pred
#share 5% of prey

ShPrey <- pred.by.prey %*% prey.by.pred
cs <- ShPrey/colSums(prey.by.pred)
cs[cs<0.25] <- 0
cs[cs>=0.25] <- 1
cs[is.na(cs)] <- 0 
ShPrey <- t(cs)
diag(ShPrey) <- 0

#number of competitors
TotalNuCompetitorsPertarget <- colSums(ShPrey)
SpWebStats[["Ncomp"]] <- TotalNuCompetitorsPertarget

# mean number of occurrences across competitor for each target
sh <- ShPrey*site.no.occur
richsh <- colSums(sh)
meanRangSizCompetitorPerTargetSp <- richsh / TotalNuCompetitorsPertarget
SpWebStats[["RScomp"]] <- meanRangSizCompetitorPerTargetSp

## apparent competitors--------------
ShPred <- prey.by.pred %*% pred.by.prey

#share 5% of prey
cs <- ShPred/colSums(pred.by.prey)
cs[cs<0.25] <- 0
cs[cs>=0.25] <- 1
cs[is.na(cs)] <- 0 
ShPred <- t(cs)
diag(ShPred) <- 0

#number of competitors
TotalNuAppCompetitorsPertarget <- colSums(ShPred)
SpWebStats[["Napcomp"]] <- TotalNuAppCompetitorsPertarget

# mean number of occurrences across app competitor for each target
sh2 <- ShPred*site.no.occur
richsh2 <- colSums(sh2)
meanRangSizAppCompetitorPerTargetSp <- richsh2 / TotalNuAppCompetitorsPertarget
SpWebStats[["RSapcomp"]] <- meanRangSizAppCompetitorPerTargetSp

save(SpWebStats,file = "SpeciesLevelWebStatisticsGloBI.Rdata")

rm(SpWebStats)

```



