---
title: "Table S1. List of model runs"
output: pdf_document
---

**List of model runs for different metawebs, biotic metrics, cross-validation runs, and distribution data.**

1. Metawebs: MW1=the potential metaweb with interactions for all life stages except eggs; M2=same as M1 except relationships are constrained such that predator-prey combinations must occur in at least one pixel together). MW-Globi=metaweb built from GLOBI data available for european vertebrates.

2. Biotic metrics: 25%=competitor richness (defined as species that share at least 25% of their prey species out of the total number of prey for that species); 90%=competitor richness (defined as species that share at least 90% of their prey species out of the total number of prey for that species); All=competitor richness (defined as species that share at least one prey); the logged version of both competitors (the 25% threshold) and prey; and a model including prey richness and total species richness (SR) in place of competitor richness.

3. Taxonomic groups: Birds, Mammals, Reptiles and Amphibians

4. Cross-validation runs: Cross-validation runs include different types of sampling for training and testing data (random sample, spatial block-longitude and spatial block-latitude). For each type of cross-validation, three sets of model were run including all variables (Full) and partial models (Part.) with only abiotic or only biotic variables.

5. Distribution data: data used in manuscript (Original) compared to GBIF distributions (only for amphibian species).

For each runs listed below, all species that had at least on prey species (all predators in the dataset) were modelled separately and two different model algorithms were used: bayesian generalized linear models and boosted regressions trees (>10k total models)

```{r setup,echo=FALSE}
knitr::opts_knit$set(root.dir = '~/Dropbox/Manuscripts/EuropeanFoodWebCons/manuscript/Manuscript/Analyses3Feb20')
```

```{r, echo=FALSE,results='hide',warning=FALSE}
library(knitr)
library(kableExtra)


mw <- c('MW1','MW4',rep('MW2',16))
met <- c('25%','All','All','25%','90%','log','SR',rep('25%',11))
pred <- c(rep("Full",7),rep(c("Full Model","Biotic Only","Abiotic Only"),3),rep("Full Model",2))
cross <- c(rep("All data",7),rep('Random',3),rep('Spatial-long',3),rep('Spatial-lat',3),rep('All data',2))
dist <- c(rep('Original',16),rep('GBIF',2))                            
     
tab <- cbind(mw,met,pred,cross,dist)

tab <- rbind(tab,tab[1:16,],tab[1:16,],tab[1:16,])

indx <- seq(1,nrow(tab),1)
gr <- c(rep("Amph.",18),rep("Bird",16),rep("Mammal",16),rep("Reptile",16))

tab <- cbind(indx,gr,tab)

colnames(tab) <- c("Set of runs","Group","Metaweb","Biotic Metric","Predictors","Cross-validation","Distribution Data")




```


```{r, echo=FALSE,warning=FALSE}

kable(tab, longtable=TRUE) %>%
kable_styling(latex_options = c("scale_down","repeat_header"))
#"scale_down"
```
                 
